---
title: 'Homework #10'
author: "Kate Mykhaylova"
date: "4/26/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(classdata)
```

```{r}
acc <- read.csv("https://raw.githubusercontent.com/DS202-at-ISU/labs/master/data/fars2016/accident.csv", stringsAsFactors = FALSE)
names(acc)
```

## Part one: Accident data

Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?<br>
```{r}
acc.day.week = acc %>% group_by(DAY_WEEK) %>% summarise(n = n())
acc.day.week = acc.day.week %>% mutate(DAY_WEEK = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
acc.day.week.r = acc.day.week$DAY_WEEK
level.reorder = levels(reorder(acc.day.week$DAY_WEEK, acc.day.week$n, fun = mean, na.rm=TRUE))
acc.day.week.reorder = factor(acc.day.week.r, levels = level.reorder)
acc.day.week$DAY_WEEK = acc.day.week.reorder
acc.day.week %>% ggplot(aes(x = DAY_WEEK)) +
  geom_bar(aes(weight = n))
# There are more accidents occuring on Friday, Saturday, and Sunday, with most happening on Saturday.
```

What time of the day do accidents happen (use variable HOUR)?<br>
```{r}
acc.hour = acc %>% group_by(HOUR) %>% summarise(n = n())
acc.hour <- acc.hour[-c(25), ]
acc.hour %>% ggplot(aes(x = HOUR)) +
  geom_bar(aes(weight = n))
# Most accidents happen at 6pm.
```

What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?<br>
```{r}
acc.drunk = acc %>% filter(DRUNK_DR >= 1) %>% summarise(n = n())
acc.drunk
# There are 8720 accidents with at least one drunk driver.
```

## Part two: Connecting data

Connect to the [person table](https://raw.githubusercontent.com/yumouqiu/DS202-Spring2022/main/Practice/data/fars2016/person.csv). Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.
```{r}
# person.table = read.csv("https://github.com/yumouqiu/DS202-Spring2023/blob/main/Practice/data/fars2016/person.csv")
# person.table = person.table %>% filter(PER_TYP == 1)
# error when trying to read this data, can't find the object PER_TYP
```

Join accident and driver table (work out which variable(s) to use)
```{r}
# person.table = person.table %>% filter(SEX == 1 | SEX == 2)
# acc.person = left_join(acc, person.table, by = c("ST_CASE", "HOUR", "MINUTE", "DAY", "MONTH"))
```

Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!
```{r}
# acc.person.tally = acc.person %>% group_by(DAY_WEEK, HOUR, SEX) %>% 
#  summarise(n = n())
# acc.person.tally = na.omit(acc.person.tally)
# acc.person.tally = acc.person.tally %>% filter(HOUR < 25)
# acc.person.tally %>% ggplot(aes(x = HOUR, y = n, color = SEX)) +
#  geom_point() + 
#  facet_wrap(~DAY_WEEK)
```

