---
title: "Homework 5"
author: "Kate Mykhaylova"
date: "3/6/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(classdata)
library(tidyverse)
library(ggplot2)
library(plyr)
```

Loading in the Data
```{r}
choco = read.csv("https://ds202-at-isu.github.io/labs/data/choco.csv")
```

## Question Set 1

1. What is the overall number of chocolate bars rated?
```{r}
str(choco)

#Total chocolate bars rated is 1852
```
2. How does the number of ratings depend on the year? Draw a bar chart of the number of reports
```{r}
choco %>% ggplot(aes(x = as.factor(Review.Date))) + 
  geom_bar() +
  ggtitle("") +
  xlab("Year")

# More chocolate bars got rated as the years went on until 2015 where the number peaked. 2016 was lower and 2017 was even lower. While the rise from 2006 to 2015 was not exactly linear it still looks like it has a big enough correlation to consider that there is an linera trend up until the peak.
```

## Question Set 2

1. How are ratings distributed? Draw a histogram of ratings and describe it. Don’t forget to mention outlines, if there are any.
```{r}
choco %>% ggplot(aes(x = as.factor(Rating))) + 
  geom_bar() +
  ggtitle("") +
  xlab("Year")

boxplot.stats(choco$Rating)$out

# The ratings seem to follow a simple normal distribution centered at 3.25 and extends to 2.5 and 4. I do not think that any rating below 2.5 or above 4 could be considered part of this distribution. Anything under or equal to 2 and any 5 value is considered to be an outlier.
```

2. Do ratings depend on the cocoa percentage of a chocolate bar? <br>
```{r}
choco %>% ggplot(aes(x = Cocoa.Pct, y = Rating)) + 
  geom_point() 

data1 <- choco[, c(5, 7)]
round(cor(data1), 2)

# There is no correlation between rating and cocoa percentage since the correlation value is -0.16.
```

3. How do ratings compare across different company locations? Focus on the three locations with the most ratings: <br>
```{r}
choco1 <- dplyr::filter(choco, Company.Location %in% c("U.S.A.","France", "Canada"))
choco2 <- dplyr::filter(choco, Company.Location %in% c("U.S.A."))
choco3 <- dplyr::filter(choco, Company.Location %in% c("Canada"))
choco4 <- dplyr::filter(choco, Company.Location %in% c("France"))

ggplot(data = choco1, aes(x = Company.Location)) +
  geom_bar(aes(weight = Rating), position = position_dodge()) +
  facet_wrap(~Rating)

str(choco2) # USA  - 787
str(choco3) # Canada - 132
str(choco4) # France - 158

mean(choco2$Rating)
mean(choco3$Rating)
mean(choco4$Rating)

# As we can see, America makes the most chocolate with Canada and France far behind. However, the average rating for Canada and France is higher than for America. The plot shows America's amount being more than double Canada and France combined, but the avergae being worse than both.
```


## Your Own Question

1.  On average, which country has the highest rated chocolate?
```{r}
data2 <- ddply(choco, .(Company.Location), function(x) mean(x$Rating))
names(data2)[2] <- 'AverageRating'

data2 %>% ggplot(aes(x = Company.Location)) +
  geom_bar(aes(weight = AverageRating)) +
  coord_flip()

data2[which.max(data2$AverageRating),]

#From the chart we can see that Chile has the highest rated chocolate on average with value of 3.75. We can also see that the average rating is consistent across countries at ~3.
```

## Summary of Workflow
Prefer to work alone, worked on this project in my own time.


